---
title: "Development of restored dike grasslands to reference needs more than four years and can hardly modified by local site conditions"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  officedown::rdocx_document:
    reference_docx: StylesTemplate.docx
  #bookdown::word_document2:
csl: template-journal-of-vegetation-science.csl
bibliography: bibliography.bib
---

```{r set-up, include = FALSE}
library(knitr)
library(bookdown)
library(officedown)
library(officer)
library(tidyverse)
library(flextable)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  dpi = 300,
  fig.cap = FALSE
  )
set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11
  )
```

# Document head <!--# Remove --> {.unnumbered}

Restoration Ecology, TUM School of Life Sciences, Technical University of Munich, Germany

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

Max 8000 words <!--# Remove -->

\newpage

# Abstract (max 300 words) {.unnumbered}

\clearpage

\newpage

# Keywords {.unnumbered}

Bayesian models

dike grasslands

embankment

endangered species

environmental filter

species composition

restoration

sand

soil properties

surveys

vegetation

\clearpage

\newpage

# Introduction

"Ecological theory can point restoration toward important processes that need manipulation [...]. However, for this information to be relevant, restoration ecology needs to employ evidence-based assessments [...]" (Suding 2011) --\> No evidence for theory with our project

-   Endangered grasslands

-   Predictability of restorations --\> experiments, evidence-based restoration (cooke 2018)

-   Brudvig 2017 among restoration outcomes and in comparison to reference

-   Substrate modification to influence habitat filtering

-   Sowing a typical approach + priority effects + species pool

-   Introduce FCS, Recovery time, Persistence

-   Dike grasslands as case study

-   The aim of the study is

-   <!--# We measured FCS, recovery, persistence -->

We asked following question to evaluate the restoration success after four years:

1.  How strong differs the persistence of the sown species among the seed mixture-substrate combinations?

2.  How strong differs the Favourable Conservation Status (FCS) among the seed mixture-substrate combinations?

3.  How close is the vegetation to the reference state?

\clearpage

\newpage

# Material & Methods

## Experimental design {.unnumbered}

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:figure-map)); 314 m above sea level; WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and an annual precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@hari2020; @cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [Appendix Table \@ref(tab:flood-weather), @lfu2021waterlevel; @lfu2021discharge]. Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from the dike<!--# Is that right?  --> and a with an excavator the substrates were mixed <!--# Is that right?  --> and the plots prepared. The target vegetation were lowland hay meadows (Arrhenatherion elatioris, CM01A) and calcareous grassland [Cirsio-Brachypodion pinnati, DA01B code of the EuroVegChecklist, @mucina2016]. The species pool of hay meadows and calcareous grasslands consisted of 55 and 58 species , respectively. The seeds were received from a commercial producer of autochtonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(tab:traits)) . The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, Bromus hordeaceus was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition, halfway up the slopes of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-map)). The split plot was caused by testing all 24 treatments on both sides of the dike. We tested three substrates with 0%, 25% and 50% sand admixture and two soil depths (15 vs. 30 cm). Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. We tested soil samples of the three substrates on both expositions by mixing several subsamples. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table \@ref(tab:substrate)). The two seed mixture types were tested with two seeding densities (4 vs. 8 g m^-2^). The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. Establishment rates of species (Appendix Table \@ref(fig:table-establishment)).

## External data

Reference vegetation surveys were extracted from sPlotOpen [@Sabatini.2021] and our own surveys in the surroundings on Danube dikes [@bauer2023]. We selected six calcareous grassland plots [EUNIS code R1A, @Chytry.2020] within SE Germany from sPlotOpen and 98 plots of our own survey which included furthermore lowland hay meadows (R22) and as a negative reference ruderal, dry and anthropogenic vegetation (V38).

<!--# Mention traits -->

## Statistical analyses {.unnumbered}

To measure the effects of the treatments on our three response variables, we calculated Bayesian linear mixed effects models (BLMM) with the random effect plot nested in block with the Cauchy prior [see @lemoine2019]. Furthermore, we included as a fixed effect the botanists of each plot. For the simple effects of the treatments, we chose plausible weakly informative priors. To evaluate the influence of the priors, prior predictive checks and models with non-informative priors were calculated. For the computation, we used 4 chains, a thinning rate of 2, 5,000 iterations for warm-up and 10,000 in total. For evaluating the computation, MCMC chain convergence was checked with trace plots and by the R-hat values and MCMC chain resolution by the effective sampling size (ESS). Posterior predictive check was done with Kernel density estimates histograms of statistics skew and leave-one-out (LOO) cross-validation [see @gabry2019]. Finally, models were compared with the Bayes factor (BF) and Bayesian *R*² values [@gelman2019].

A non-metric multidimensional scaling ordination (NMDS) with Jaccard dissimilarity was used to visualize variation in species composition in space and time. Seven species which had an accumulated cover over all plots of less than 0.5% were excluded (2D-stress: 0.21).

Data, code and the whole model specification and evaluation is stored on Zenodo and in an easily accessible a scroll-down document on GitHub [cite Zenoto]<!--# Cite Zenodo--> There, the sections are referenced to the Bayesian analysis reporting guidelines [BARG, @kruschke2021]. All analyses were performed in R [Version 4.2.2, @base] with the functions 'brm' from the package 'brms' [@brms] for model calculation, several functions from 'brms' and 'bayesplot' [@bayesplot] for model evaluation and 'metaMDS' from 'vegan' for the ordination [@vegan].

\clearpage

\newpage

# Results

We could identify a statistically clear positive effect of sand addition on the persistence of sown species and on the recovery rate (Figure \@ref(fig:figure-effects)). Furthermore, sown hay meadows were clearly closer to their reference plots than sown dry grasslands were to their references. The posterior distributions which are shown in the interactions plots which separates exposition and survey year indicate for all three responses a strong positive development after one year with slow increase in following years (Figures \@ref(fig:figure-persistence); \@ref(fig:figure-fcs); \@ref(fig:figure-recovery)) . Both expositions have similar trends but for all responses the values are clearly lower on the south exposition. Persistence values are more than 50% higher on the north exposition (Figure \@ref(fig:figure-persistence)). The effect of restoration treatments is in the interaction plots not clear. Hay meadow seed mixture led to a closer development to hay meadow references and almost reached on the north exposition the area of the standard deviation of reference sites (Figure \@ref(fig:figure-recovery)). Though, the sand admixture had no clear effect on the development. Persistence of both seed mixture types seems to be slightly positive affected by sand admixture on the north exposition (Figure \@ref(fig:figure-persistence)).

The ordination visualizes the species composition of seed mixtures and the development of the plots over time (Figure \@ref(fig:figure-nmds)). The NMDS confirms that the seed mixtures are variable, but distinctive for hay meadows and dry grasslands. In addition, it visualizes the decreasing recovery rate on the NMDS1 axis which was already presented (Figure \@ref(fig:figure-recovery)) but the right direction of the development. One exception is that the development of intended hay meadows on the south exposition does not develop towards its seed mixture compositions. The ordination presents the experimental plots not only in context with seed mixtures, but also with reference sites. The reference sites have a bigger variation than the seed mixtures and are close to the seed mixtures but hardly overlap. The positions of reference sites is shifted to the left on the NMDS1 axis which means in the direction of early successional states. On the south exposition, we had negative reference sites which are located in the NMDS in between the positive reference sites and the recent state of restored plots. However, most of negative reference sites (ruderal, dry and anthropogenic vegetation) are situated in the upper half and explicitly recent restoration sites on the south exposition are close to them.

\clearpage

\newpage

# Discussion

Mention drought in 2018 and 2019

\clearpage

\newpage

# Conclusions

Text

\clearpage

\newpage

# Acknowledgements {.unnumbered}

We would like to thank our project partners Dr. Markus Fischer, Frank Schuster, and Christoph Schwahn (WIGES GmbH) and Klaus Rachl and Stefan Radlmair (Regierung von Niederbayern) for numerous discussions on restoration and management of dike grasslands. Field work was supported by Clemens Berger and Uwe Kleber-Lerchbaumer (Wasserwirtschaftsamt Deggendorf). We thank Holger Paetsch, Simon Reith, Anna Ritter, Jakob Strak, Leonardo H. Teixeira, and Linda Weggler for assisting with the field surveys or soil analyses in 2018--2020. The German Federal Environmental Foundation (DBU) supported MB with a doctoral scholarship.

# Author contribution {.unnumbered}

JH and JK designed the experiment. JH did the surveys in the years 2018--2020 and MB in 2019 and 2021. MB did the analyses and wrote the manuscript. JK and JH critically revised the manuscript.

# Open research {#sec-open-research .unnumbered}

Data and code is stored on Zenodo (Bauer et al. 2023) <!--# Update citation after upload -->

Model evaluation is stored on GitHub: [github.com/markus1bauer/2023_danube_dike_experiment](https://github.com/markus1bauer/2023_danube_dike_experiment/tree/main/markdown)

# Funding {.unnumbered}

MB is funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020. <!--# funding number of WIGES -->

# References {.unnumbered}

::: {#refs}
:::

\clearpage

\newpage

<!---BLOCK_LANDSCAPE_START--->

# Tables {.unnumbered}

## Table 1 {.unnumbered}

Each plot received an individual set of 20 species with some restrictions to the number of species per functional group. The total species pool for hay meadows was 55 and for calcareous grassland 58. In total, seed mixtures had 16 or 32 g per plot.

```{r table-seedmix, tab.id = "seedmix", tab.cap = "Seedmix"}
read_csv(here::here("data", "raw", "data_raw_table_1_seedmix.csv"),
         show_col_types = FALSE) %>%
  slice(-1) %>%
  flextable() %>%
  merge_at(part = "header", j = 2:3) %>%
  align(align = "center", part = "all") %>%
  align(align = "left", part = "all", j = 1) %>%
  add_header_row(top = FALSE, values = c("", "Hay meadow", "Calcareous grassland", "", "", "")) %>%
  add_header_row(top = FALSE, values = c("", "#", "#", "#", "%", "%")) %>%
  bold(part = "header", i = 1) %>%
  autofit()
```

\clearpage

\newpage

## Table 2 {.unnumbered}

functional traits <!--# full description -->

```{r table-traits, tab.id = "traits", tab.cap = "Traits"}
#read_csv(here::here("data", "raw", "data_raw_table_3.csv")) %>%
  #flextable() %>%
  #bold(part = "header") %>%
  #autofit()
```

\clearpage

\newpage

## Table 3 {.unnumbered}

Soil of the three substrates was analyzed for the fraction \< 2mm fraction. Soil texture was classified according to the *Bodenkundliche Kartieranleitung* [@BodenkundlKartieranl.2005]. The pH was measured in CaCl~2~ solution. Plant available phosphorus and potassium were measured in a calcium acetate-lactate extract and magnesium in CaCl~2~ extract. For calculating CaCO~3~, a subsample was annealed at 550 °C and the measured C amount calculated with 8.33. To calculate total N and the C/N ratio, a subsample was incinerated at 1000 °C.

```{r table-substrate, tab.id = "substrate", tab.cap = "Substrate"}
read_csv(here::here("data", "raw", "data_raw_table_3_substrate.csv"),
         show_col_types = FALSE) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(
      here::here("data", "raw", "data_raw_table_3_substrate.csv"),
                 show_col_types = FALSE) %>%
      slice(1)
    ) %>%
  align(align = "center", part = "all") %>%
  align(align = "left", part = "all", j = 1) %>%
  compose(i = 1, j = 10, part = "header", value = as_paragraph("P", as_sub("2"), "O", as_sub("5"))) %>%
  compose(i = 1, j = 11, part = "header", value = as_paragraph("K", as_sub("2"), "O")) %>%
  compose(i = 1, j = 12, part = "header", value = as_paragraph("Mg", as_sup("2+"))) %>%
  compose(i = 1, j = 14, part = "header", value = as_paragraph("CaCO", as_sub("3"))) %>%
  compose(i = 2, j = 10, part = "header", value = as_paragraph("mg 100g", as_sup("−1"))) %>%
  compose(i = 2, j = 11, part = "header", value = as_paragraph("mg 100g", as_sup("−1"))) %>%
  compose(i = 2, j = 12, part = "header", value = as_paragraph("mg 100g", as_sup("−1"))) %>%
  set_table_properties(layout = "autofit", width = 1)
```

\newpage

<!---BLOCK_LANDSCAPE_STOP--->

# Figures {.unnumbered}

## Figure 1 {.unnumbered}

```{r figure-map, fig.cap = "Map", fig.asp = 0.7272}
knitr::include_graphics(here::here("outputs", "figures", "figure_1_map_tmap_300dpi_8x11cm.tiff"))
```

The experiment was located

\clearpage

\newpage

## Figure 2 {.unnumbered}

```{r figure-effects, fig.cap = "Effects", fig.asp = 3.3}
knitr::include_graphics(here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff"))
```

\clearpage

\newpage

## Figure 3 {.unnumbered}

```{r figure-persistence, fig.cap = "Persistence", fig.asp = 1.785}
knitr::include_graphics(here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x14cm.tiff"))
```

\clearpage

\newpage

## Figure 4 {.unnumbered}

```{r figure-fcs, fig.cap = "Favourable Conservation Status (FCS)", fig.asp = 1.785}
knitr::include_graphics(here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x14cm.tiff"))
```

\clearpage

\newpage

## Figure 5 {.unnumbered}

```{r figure-recovery, fig.cap = "Recovery", fig.asp = 1.785}
knitr::include_graphics(here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x14cm.tiff"))
```

Fig. 5:

\clearpage

\newpage

## Figure 6 {.unnumbered}

```{r figure-nmds, fig.cap = "NMDS", fig.asp = 1.0312}
knitr::include_graphics(here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff"))
```

Fig. 6:

\clearpage

\newpage

<!---BLOCK_LANDSCAPE_START--->

# Supplementary Material {.unnumbered}

## Table S1 {.unnumbered}

Three dry years (2018--2020) and three minor floods (2018, 2019, 2021) occurred during the study period. Annual temperature, precipitation and discharge of River Danube near the study sites (2017--2021), based on weather station Metten [mean, 1981--2010; ID, 3271; WGS84, lat/lon, 48.85476, 12.918911; @cdc2022prec, @cdc2022temp] and stream gauge Pfelling [ID, 10078000; WGS84, lat/lon, 48.87975, 12.74716; @lfu2021discharge]. HQ2 = Highest discharge with a probability of occurrence every second year. HSW = Highest water level for shipping.

```{r table-flood-weather, tab.id = "flood-weather", tab.cap = "Flood and weather data"}
read_csv(here::here("data", "raw", "data_raw_table_s1.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  compose(part = "body", i = ~ Unit == "m3 s−1", j = "Unit", value = as_paragraph("m", as_sup("3"), " s", as_sup("−1"))) %>%
  autofit()
```

\clearpage

\newpage

<!---BLOCK_LANDSCAPE_STOP--->

## Table S2 {.unnumbered}

Establishment rate <!--# Full description an include it here -->

```{r table-establishment, fig.cap = "table-s2"}
#knitr::include_graphics(here::here("outputs", "tables", "table_a2_establishment.html"))
```

\clearpage

\newpage

## Table S3 {.unnumbered}

See repository outputs/tables/table_a3_seedmixes.csv <!--# Write full link and full description -->

\clearpage

\newpage

## Session Info {.unnumbered}

```{r sessioninfo}
sessionInfo()
```
