---
title: "Fit by design: Developing substrate-specific seed mixtures for functional dike restoration"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  officedown::rdocx_document:
    reference_docx: template_word.docx
csl: template_references_j_appl_ecol.csl
bibliography: bibliography.bib
---

```{r set-up, include = FALSE, message = FALSE}
library(here)
library(knitr)
library(officedown)
library(officer)
library(tidyverse)
library(flextable)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  dpi = 300,
  fig.cap = FALSE
  )
set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  text.align = "center"
  )
ft <- officer::fp_text(shading.color = "yellow")

word_spec <- function(x, prop = ft) ftext(text = toString(x) ,prop =  ft)
```

Restoration Ecology, TUM School of Life Sciences, Technical University of Munich, Germany

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

\clearpage

\newpage

# Abstract {.unnumbered}

1.  Sowing is a well-established restoration technique to overcome dispersal limitation and site-specific seed mixtures are most effective to achieve functional communities. This is especially important if the restored vegetation has to protect critical infrastructure like roadsides and dikes. Here, an improved seed--substrate combination would secure slope stability, reduce mowing efforts, and generate biodiverse grasslands.
2.  A factorial field experiment addressed this topic on a dike at river Danube in S Germany in the period 2018--2021. On 288 plots, we tested three sand admixtures, two substrate depths, two seed densities and two seed mixtures (mesic hay meadow, calcareous grassland) in north or south exposition, and measured the persistence of sown species, the Favorable Conservation Status (FCS), and the successional distance to reference sites as calculated from an ordination.
3.  Sand admixture had a slightly positive effect on target variables, while substrate depth, seed density and mixture type had no effects on species persistence or FCS.
4.  Overall, the sown vegetation developed in the desired direction, but a recovery debt remained after four years, and some plots still showed similarities to negative references from ruderal sites. In north exposition, hay meadow-seed mixtures developed closer to the respective reference communities than dry grasslands.
5.  In south exposition, the sown communities developed poorly which might be due to a severe drought during establishment and the addition and removal of a geotextile. This initial negative effect remained over the entire observation period.
6.  Synthesis and applications: Site-adapted seed mixtures make restoration more efficient. However, applying several seed--substrate combinations might foster beta diversity. Furthermore, additional efforts in management might be necessary to reduce the recovery debt and re-sowing after unfavorable conditions is recommended.

\clearpage

\newpage

# Keywords {.unnumbered}

artificial soil mixture

community assembly

dry grasslands

ecological restoration

Favorable Conservation Status

hay meadow

levee

persistence

river embankment

soil restoration

sowing

species composition

species-rich grassland

\clearpage

\newpage

# Introduction

Grasslands can support a very high biodiversity and they provide several ecosystem services [@Dengler.2014; @Bardgett.2021]. However, they are globally endangered [@Bardgett.2021] for example by misguided afforestation [@dudley2020], and in Europe, calcareous grasslands and hay meadows are red-listed habitats [Category 3, 'vulnerable', @Janssen.2016]. Restoration is seen as a key factor to sustain biodiversity and ecosystem services [@COP.2014; @UnitedNations.2019b], and sowing is a well-established approach to establish species-rich grasslands [@Kiehl.2010]. Sowing high-diversity mixtures of local provenance produced by specialized companies is a promising way to scale up restoration efforts [@Freitag.2021], and to overcome the dispersal filters [@myers2009; @orrock2023]. However, there are still open questions about adjusting the seed mixtures to specific site conditions and future climate conditions [@Torok.2021].

Restoration ecology can increase predictability of restoration approaches [@Suding.2011; @Mouquet.2015] by using rigorous, repeatable, and transparent experiments based on advanced theory, which will finally lead to evidence-based restoration [@Cooke.2018; @Wainwright.2018]. Local site conditions and the restoration method are key predictors of the development of vegetation after sowing [@Brudvig.2017b] and habitat and biotic filtering are the main assembly factors which can be manipulated by the choice of the seed--substrate combination [@Torok.2017]. This means a close adaptation of the substrate to the niche of the target species or of the seed mixtures to the characteristics of the chosen substrate. Suitable substrates should reduce habitat filtering, while the seed mixture should minimize competitive exclusion of unsuitable species. Modifying seed mixtures to match specific site conditions could be based on functional plant traits [@Funk.2008; @Laughlin.2014; @Balazs.2020], although this is not easy to implement [@merchant2022; @Bauer.2022]. This challenge is particularly interesting on artificial substrates that are used in urban areas [@Bauer.2022], quarries [@chenot-lescure2022], or on dikes [@Liebrand.1996].

Dikes are promising sites for the restoration of species-rich grasslands because they can increase habitat area and connectivity of semi-natural grasslands and therefore significantly contribute to biodiversity conservation in agricultural landscapes [@Batori.2020]. Steep slopes with different exposition, contrasting substrate layers and dense swards for erosion protection characterize dikes [@Batori.2016; @berendse2015]. Dikes can reconcile several requirements [@teixeira2022], which can be fostered by an adapted seed--substrate combination.

The aim of this study is to identify the best combination of seed mixtures and substrates for vital and species-rich grasslands on north- and south-exposed dike slopes. An experiment was set up to test agricultural soil with different amounts of sand admixture, substrate depths, seed densities, and seed mixture types. We expected a better development of calcareous grassland in south exposition with shallow sandy substrates, and mesic hay meadows in north exposition on less sandy deep substrate. Furthermore, higher seed densities were recommended for slopes [@KleberLerchbaumer.2017], albeit without experimental evidence.

To measure the success of restoration, we chose four indices. First, we observed the persistence of the sown species, which is a key objective of this restoration measure [@Wilsey.2021]. For this purpose, presence of the sown species was monitored for three consecutive years. Second, we calculated the Favorable Conservation Status (FCS) which distinguishes characteristic diversity and derived diversity [@Helm.2015]. Characteristic diversity is created by established species which belong to the habitat-specific species pool. In other words, these species are associated with the habitat type of interest. Third, the successional distance to the reference grasslands was determined. This measure quantifies the ordination distance of a plot to the average reference site [@Rydgren.2019]. To complement these indices, the success of restoration or the difference from desired conditions was evaluated [cf. @Brudvig.2017b]. This was reached by comparing the species composition of the experimental plots with reference sites. We asked the following question to evaluate the restoration approaches and the success after four years:

1.  How strong differs the persistence of the sown species among the seed--substrate combinations?

2.  How strong differs the Favorable Conservation Status (FCS) among the seed--substrate combinations?

3.  How close is the established vegetation to the reference state?

\clearpage

\newpage

# Material & Methods

## Experimental design {.unnumbered}

The combinations of seed mixtures and substrates ('seed--substrate combinations') were tested on a dike at river Danube in SE Germany (Figure \@ref(fig:map); 314 m a.s.l.; WGS84 (lat/lon), 48.83895, 12.88412). The climate was temperate-suboceanic with a mean annual temperature of 8.4 °C and an annual precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@hari2020; @cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [Appendix A1, @lfu2021waterlevel; @lfu2021discharge]. The substrates consisted of sand (0--4 mm) and agricultural soil obtained from a nearby dike construction site near the village of Steinkirchen. A big roller mixed both components and an excavator put the substrates in the prepared plots.

The target vegetation were lowland mesic hay meadows and semi-dry calcareous grassland [EUNIS code: R22, R1A, @Chytry.2020, 'Arrhenatherion elatioris' and 'Cirsio-Brachypodion pinnati' according to the EuroVegChecklist: CM01A, DA01B, @mucina2016]. The species pool of hay meadows and dry grasslands consisted of 55 and 58 species, respectively. The seeds were received from a commercial producer of autochthonous seeds (Co. Krimmer, Pulling). From these species pools, 20 species were randomly selected for each plot (Appendix A2). The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Appendix A3). As expected, the hay-meadow mixtures had a higher specific leaf area (SLA), lower seed mass, and higher canopy height than the dry-grassland mixtures. The south-exposed plots were sown in mid-April 2018 and the north exposition 14 days later. Due to the drought, the south exposition was protected by a geotextile consisting of straw chaff (350 g m^-2^) which was removed after two weeks due to unsatisfactory effects like the death of several protected lizards. The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by standard deep cuts with hay removal in July 2019 and 2020. In October 2018, *Bromus hordeaceus* was sown as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were halfway up the slopes of the dike, distributed over the north and south exposition, and arranged in six blocks (= replicates). The experiment used a split-plot design combined with a randomized complete block design (Figure \@ref(fig:map)). The split plot was created by the two expositions of the dike, where all 24 treatment combinations were tested, i.e., sand admixtures (0, 25, and 50%), soil depths (15 vs. 30 cm), and the two seed mixture types with the normal seed density and an increased density recommended for slopes [4 vs. 8 g m^-2^, @Kiehl.2010; @KleberLerchbaumer.2017].

Under the substrate, a 5 cm thick drainage layer of gravel (0-16 mm) was installed. Soil samples of the three substrates were tested on both expositions by mixing several sub-samples from different plots. The sand admixture changed the soil texture, increased the C/N ratio, reduced Calciumcarbonat, but did hardly changed the quite high pH values (Table \@ref(tab:substrate)). Phosphate and potassium were rather scarce for agricultural soils and magnesium high [@lfl2022]. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. The establishment rates of species are recorded in Appendix A4. In total, 48 species of the species pool of hay meadows (87%) and 46 (79%) of dry grasslands) established by 2021. The species established 31 ± 22% of their sown plot. In total, 274 species were detected (Appendix A5).

Reference vegetation surveys were extracted from sPlotOpen [@Sabatini.2021] and our own surveys on Danube dikes in the surroundings [@bauer2023]. We selected six calcareous grassland plots [EUNIS code R1A, @Chytry.2020] within SE Germany from sPlotOpen and 98 plots of our own survey which included also lowland hay meadows (R22) and as a negative reference ruderal, dry and anthropogenic vegetation (V38).

## Statistical analyses {.unnumbered}

To measure the effects of the treatments on our three response variables, we calculated Bayesian linear mixed-effects models (BLMM) with the random effect plot nested in block with the Cauchy prior [see @lemoine2019]. Furthermore, we included as a fixed effect the botanists recording a certain plot. For the simple effects of the treatments, we chose plausible weakly informative priors. To evaluate the influence of the priors, prior predictive checks and models with non-informative priors were calculated. For the computation, we used four chains, a thinning rate of two, 5,000 iterations for warm-up, and 10,000 in total. For evaluating the computation, MCMC chain convergence was checked with trace plots and by the R-hat values, and MCMC chain resolution by the effective sampling size (ESS). Posterior predictive checks were done with Kernel density estimates histograms of statistics skew and leave-one-out (LOO) cross-validation [see @gabry2019]. Finally, models were compared with the Bayes factor (BF) and Bayesian *R*² values [@gelman2019]. A non-metric multidimensional scaling ordination (NMDS) with Jaccard dissimilarity was used to visualize variation in species composition in space and time (2D-stress 0.21). Seven species were excluded because they had an accumulated cover over all plots of less than 0.5%. In total, 343 species were included in the ordination.

Data, code and the entire model specifications and evaluations are stored on GitHub and presented in an easily accessible document for scrolling through `r word_spec("(include Zenodo reference)")`.<!--# Cite Zenodo--> There, the sections are referenced to the Bayesian analysis reporting guidelines [BARG, @kruschke2021]. All analyses were performed in R [Version 4.2.2, @base] with the functions 'brm' from the package 'brms' [@brms] for model calculation, several functions from 'brms' and 'bayesplot' [@bayesplot] for model evaluation, and 'metaMDS' from 'vegan' for the ordination [@vegan].

\clearpage

\newpage

# Results

## Weak effects of restoration treatments

We could identify a statistically clear positive effect of sand admixture on the persistence of sown species and on the recovery rate (Figure \@ref(fig:effects)). The posterior distributions are also shown in the interaction plots that separate exposition and survey year (Figures \@ref(fig:persistence), \@ref(fig:fcs), \@ref(fig:recovery)). For all three response variables, the vegetation developed strongly positive after one year, but the recovery rate slowed down in the following years. Both expositions had similar trends but for all responses the values were clearly lower in south exposition, e.g., persistence values were on average more than 46% higher in north exposition (Figure \@ref(fig:persistence)). For the interactions, the effects of restoration treatments were neither clear nor strong (Figures \@ref(fig:persistence), \@ref(fig:fcs), \@ref(fig:recovery)). Persistence of both seed mixture types was slightly positively affected by sand admixture on the north exposition (+ 6--7 ± 4%, Figure \@ref(fig:persistence)).

## Hay meadows on north exposition closer to reference

The ordination visualizes the species composition of seed mixtures and the development of the plots over time (Figure \@ref(fig:nmds)). The NMDS confirmed that the seed mixtures were variable, but distinctive for hay meadows and dry grasslands and ascertained the intended direction of the vegetation development. One exception was that the development of hay meadows in south exposition did not develop towards its seed-mixture compositions.

The ordination presents the experimental plots also in context with reference sites (Figure \@ref(fig:nmds)). The reference sites had a larger variation than the seed mixtures and were close to the seed mixtures but hardly overlapped. The positions of reference sites were shifted to the left in comparison to the seed mixtures, which means in the direction of early successional states. Though, they were still different from the negative references of ruderal vegetation. Negative references were only available on the south exposition and they were located in the NMDS between the positive reference sites and the state of restored plots in 2021. Hay meadow-seed mixtures led to a closer development to hay meadow references than dry grasslands to their references (Figure \@ref(fig:effects), \@ref(fig:recovery)) this was especially the case on the north exposition (Figure \@ref(fig:nmds)).

\clearpage

\newpage

# Discussion

The vegetation of the dike experiment developed in the desired direction, but with a decelerating recovery rate. On the south exposition, a significant worse effect during establishment phase remained over the complete observation period. All in all, it was hardly possible to identify an best seed--substrate combination. Though, hay meadows developed closer to its reference on the north exposition than dry grasslands.

## General effects of treatments and exposition

Restoration on agricultural soils can have limited success due to high nutrient loads [@walker2004] but can lead to positive results, if mixed with mineral material as sand and gravel [@chenot-lescure2022]. As expected, sand admixture reduced nutrient loads and led to higher persistence of sown species while 50% admixture did not further increase the effect. In addition, the effect only appeared in north exposition and the effect size of about 6% in the fourth year after restoration was rather small. Substrate depth did not significantly affect persistence or Favorable Conservation Status (FCS) similarly to earlier studies [@baer2004]. Larger differences in soil depths might be necessary to observe negative effects by thicker substrate layers as was shown for prairies [@dornbush2010]. Neither, seed density had an effect on persistence and FCS which fits to results of @Kaulfuss.2022, who found out that a certain amount of seeds are necessary for a successful establishment of target species but higher densities do not further improve the result and have rather a slight negative effect.

The vegetation in south exposition had a different species composition which confirms the findings of @Batori.2016 in Hungary. Here, the differences might have some unwanted experimental reasons: the geotextile, which had been implemented on the southern slope, was removed after two weeks. This led to the drying of many seedlings which was amplified by the intense drought of 2018 and 2019 [@hari2020]. This drought could have decreased germination and establishment of sown species also in general because a moist spring with a mesic temperature improves establishment [@Larson.2021; @orrock2023]. The interesting point here is the lasting negative effect on persistence and FCS which suggests a legacy effect by adverse weather conditions after sowing as it was observed by other studies [@Stuble.2017; @Groves.2020]. The problems during the establishment phase might have led to an alternative trajectory `r word_spec("(reference?)")`<!--# Add ref--> and could probably only be met by re-sowing. The drying of seedlings might also have leveled the distinction of the seed mixture types on the south exposition.

## No interaction effect of seed--substrate combinations

Our aim was to identify the best seed--substrate combination in respect of restoration effectiveness and biodiversity. For evaluating effectiveness, we measured the persistence of the sown species and for fostering plant biodiversity in general, we measured the FCS. Though, we could not identify an interaction effect for neither of the indices. We would have expected a better performance of hay-meadow seed mixtures with lower sand admixture and for dry grasslands with higher sand admixture. Our results suggest that, at least after four years, the substrate conditions are within the range of both seed mixture types (hay meadows, dry grasslands). Although, both types are clearly plant-sociologically and functionally distinct, they are still relative closely related why they contain shared species and could develop on similar site conditions [Appendix A3, @Oberdorfer.1993; @Oberdorfer.1993b]. Other grassland studies could identify more or less clear interactions of opposing habitat preferences or functional traits along the gradients of productivity, moisture, and nutrients [@Zirbel.2020; @Freitag.2021; @Kaulfuss.2022]. However, these studies did not work with an experimental set up of different seed--substrate combinations, but analyzed the result of habitat and biotic filtering after one, five and about 15 years, respectively. Furthermore, the non-existence of an ideal combination could be explained by priority effects which means that the species of the non-perfect seed mixture could establish earlier and preempted the available niches for the species of related habitat types [@Fukami.2015].

## The success of the restoration approaches

The seed mixtures and positive reference sites were similar but hardly overlapped (Figure \@ref(fig:nmds)). The position on the NMDS suggests that the seed mixture represent a later successional stage than the references. The NMDS shows a slightly better adaption of hay meadows to the north exposition than of dry grasslands (Figure \@ref(fig:recovery)). This can be expected from the requirements of hay meadows for mesic conditions which can be provided on north exposed dike slopes [@Batori.2020; @Oberdorfer.1993b]. On the southern exposition, the plots of hay meadows developed rather to dry grassland references which indicates an inefficient restoration by using an non-adapted seed mixture.

The vegetation developed in general in the desired direction but was still distinct from positive references and seed mixtures after four years. In south exposition, the plots were rather similar to the negative reference of dry, ruderal vegetation. The gap between goal and restoration outcome was also shown for other sowing experiments or restorations [@mitchley2012; @Engst.2016; @Kaulfuss.2022] or for dike vegetation compared with semi-natural references [@Batori.2016]. This result is not surprising since the 'recovery debt' is a general phenomenon of grassland restoration [@MorenoMateos.2017; @jones2018].

\clearpage

\newpage

# Conclusions

Our results suggest that adapted seed mixtures can increase restoration efficiency by sowing hay meadows on the north exposition and not on the south exposition of dikes. Furthermore, to reduce nutrient load by sand admixture has a clear positive effect, even if the effect is small. The question which remains is, if sand admixture is the most economical restoration measure to promote diversity on dikes. Seed density seems not to be necessary to be increased on dike slopes and soil depths of 30cm are not adverse compared to 15 cm thick substrates. There was no perfect seed--substrate combination, why we conclude that a variation of seed mixture types and different substrates along restoration sections would promote biodiversity better than a single solution [@bauer2023; @Holl.2022]. The strong and lasting negative effect of drought on the south exposition might require a re-sowing. To close the recovery debt, the adaption of the management might be promising since it is a crucial factor aside of the restoration approach and the site characteristics [@Grman.2013; @tölgyesi2021]. For example, starting the already existing sheep grazing on the experimental sites will modify the disturbance regime and improve dispersal. All in all, our results support the finding that dike grasslands can promote biodiversity in agricultural landscapes [@Batori.2020]. Though, the stubborn recovery debt highlights that restored grasslands cannot substitute old-growth grasslands [@nerlekar2020; @Buisson.2022].

\clearpage

\newpage

# Acknowledgements {.unnumbered}

We would like to thank our project partners Dr. Markus Fischer, Frank Schuster, and Christoph Schwahn (WIGES GmbH) as well as Klaus Rachl and Stefan Radlmair (Lower Bavaria Government) for numerous discussions on restoration and management of dike grasslands. Field work was supported by Clemens Berger and Uwe Kleber-Lerchbaumer (Wasserwirtschaftsamt Deggendorf). We thank Holger Paetsch, Simon Reith, Anna Ritter, Jakob Strak, Leonardo H. Teixeira, and Linda Weggler for assisting with the field surveys or soil analyses in 2018--2020. The German Federal Environmental Foundation (DBU) supported MB with a doctoral scholarship.

# Author contribution {.unnumbered}

JH and JK designed the experiment. JH did the surveys in the years 2018--2020 and MB in 2019 and 2021. MB did the analyses and wrote the manuscript. JK and JH critically revised the manuscript.

# Open research {#sec-open-research .unnumbered}

Data and code are stored on Zenodo (Bauer et al. 2023) <!--# Update citation after upload -->

Model evaluation is stored on GitHub: [github.com/markus1bauer/2023_danube_dike_experiment](https://github.com/markus1bauer/2023_danube_dike_experiment/tree/main/markdown)

# Funding {.unnumbered}

MB is funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020 (No. 80 002 312). \clearpage \newpage

# References {.unnumbered}

::: {#refs}
:::

\clearpage

\newpage

<!---BLOCK_LANDSCAPE_START--->

# Tables {.unnumbered}

## Table 1 {.unnumbered}

```{r tab.id = "seedmix", ft.align = "left", tab.cap = "Each plot received an individual set of 20 species with some restrictions to the number of species per functional group. The total species pool for hay meadows was 55 and for dry grassland 58. In total, plots received 16 or 32 g of seeds. All individual seed mixtures are stored in Appendix A2."}
read_csv(here::here("data", "raw", "data_raw_table_1_seedmix.csv"),
         show_col_types = FALSE) %>%
  slice(-1) %>%
  flextable() %>%
  merge_at(part = "header", j = 2:3) %>%
  align(align = "center", part = "all", j = -1) %>%
  add_header_row(top = FALSE, values = c("", "Hay meadow", "Dry grassland", "", "", "")) %>%
  add_header_row(top = FALSE, values = c("", "#", "#", "#", "wt%", "wt%")) %>%
  bold(part = "header", i = 1) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  autofit()
```

\clearpage

\newpage

## Table 2 {.unnumbered}

Soil samples of the three substrates were analyzed for the fraction \< 2 mm. Soil texture was classified according to the 'Bodenkundliche Kartieranleitung' [@BodenkundlKartieranl.2005]. The pH was measured in CaCl2 solution. Plant available phosphorus and potassium were measured in a calcium acetate-lactate extract and magnesium in CaCl2 extract. For calculating CaCO3, a sub-sample was annealed at 550 °C and the measured C amount multiplied with 8.33. To calculate total N and the C/N ratio, a sub-sample was incinerated at 1000 °C. Lt3 = medium clayey loam; Ls4 = strong sandy loam; Sl3 = medium loamy sand; Sl4 = strong loamy sand.

```{r tab.id = "substrate", ft.align = "left", tab.cap = "Substrates"}
read_csv(here::here("data", "raw", "data_raw_table_3_substrate.csv"),
         show_col_types = FALSE) %>%
  select(-starts_with("lfl")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(
      here::here("data", "raw", "data_raw_table_3_substrate.csv"),
                 show_col_types = FALSE) %>%
      select(-starts_with("lfl")) %>%
      slice(1)
    ) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  align(align = "center", part = "all", j = -1) %>%
  compose(i = 1, j = 10, part = "header", value = as_paragraph("P", as_sub("2"), "O", as_sub("5"))) %>%
  compose(i = 1, j = 11, part = "header", value = as_paragraph("K", as_sub("2"), "O")) %>%
  compose(i = 1, j = 12, part = "header", value = as_paragraph("Mg", as_sup("2+"))) %>%
  compose(i = 1, j = 14, part = "header", value = as_paragraph("CaCO", as_sub("3"))) %>%
  compose(i = 2, j = 10, part = "header", value = as_paragraph("mg 100 g", as_sup("−1"))) %>%
  compose(i = 2, j = 11, part = "header", value = as_paragraph("mg 100 g", as_sup("−1"))) %>%
  compose(i = 2, j = 12, part = "header", value = as_paragraph("mg 100 g", as_sup("−1"))) %>%
  align(align = "left", part = "footer") %>%
  set_table_properties(layout = "autofit", width = 1)
```

\newpage

<!---BLOCK_LANDSCAPE_STOP--->

# Figures {.unnumbered}

## Figure 1 {.unnumbered}

```{r fig.id = "map", fig.asp = 1.2854, fig.width = 5, fig.cap = "Map"}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_1_map.png")
  )
```

The experiment was located on a dike at river Danube in SE Germany. The 288 plots were allocated in six blocks (white squares on upper photo) and on the north and south exposition (central photo) [both aerial photos: @bayerischevermessungsverwaltung2023]. Four treatments were conducted: sand admixture, substrate depth, seed density, and seed mixture types H and D. The western half of a block was with a shallow substrate depth and within this half the substrates had different sand admixtures. The photo on the bottom shows one block from the side after four years (Photo: Markus Bauer).

\clearpage

\newpage

## Figure 2

```{r fig.id = "effects", fig.asp = 0.3030, fig.width = 6.49, fig.cap = "The posterior density distributions (grey) of the treatments calculated over all four surveyed years. Furthermore, the median and the 66% and 95% credible intervals are shown: the persistence of sown species (A), the Favorable Conservation Status (FCS; B) and the successional distance to reference (recovery; C). The FCS is the ratio of target species and non-target species. Note that the zero lines indicate that both levels have equal values. This means for instance that hay meadows are closer to their reference than dry grasslands  (C)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff")
  )
```

\clearpage

\newpage

## Figure 3 {.unnumbered}

```{r fig.id = "persistence", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "Persistence of sown species on the dike experiment. The plots had substrates with different sand admixtures and were sown with different seed mixture types. Per plot 20 species were sown. Shown are the medians and 95 % credible intervals of the posterior distributions which were derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x8cm.tiff")
  )
```

\clearpage

\newpage

## Figure 4 {.unnumbered}

```{r fig.id = "fcs", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The Favorable Conservation Status (FCS) of a dike experiment. The plots had substrates with different sand admixtures and were sown with different seed mixture types. The zero line indicate that target and non-target species are balanced. Positive values indicate that there are more target species. Shown are the medians and 95 % credible intervals of the posterior distributions which were derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x8cm.tiff")
  )
```

\clearpage

\newpage

## Figure 5 {.unnumbered}

```{r fig.id = "recovery", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The recovery or successional distance (*d~jt,0~*) of experimental plots to its respective reference sites (hay meadows or dry calcareous grasslands). The plots had substrates with different sand admixtures and were sown with different seed mixture types. The zero lines indicate the mean position of the reference sites for each habitat type on the NMDS axis 1 (Figure 7). The grey area marks the standard deviation of the position of the reference sites. Shown are the medians and 95 % credible intervals of the posterior distributions which were derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x8cm.tiff")
  )
```

\clearpage

\newpage

## Figure 6 {.unnumbered}

```{r fig.id = "nmds", fig.asp = 0.9696, fig.width = 6.49, fig.cap = "NMDS"}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff")
  )
```

NMDS ordination based on Sørensen dissimilarity of the species composition of 288 plots on dike grasslands over four years after sowing the seed mixtures (black squares).<!--# How many observed species --> In addition, 98 positive and negative reference plots were included (filled symbols) from older dike grasslands in the surroundings [@bauer2023] and six plots from sPlotOpen [@Sabatini.2021]. The circles show the standard error of the groups. 2D stress: 0.21.

\clearpage

\newpage

# Session Info

```{r sessioninfo}
sessionInfo()
```
