---
title: "Can dike grassland restoration profit from ideal seed mixture-substrate combinations?"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  officedown::rdocx_document:
    reference_docx: template_word.docx
csl: template_references_j_appl_ecol.csl
bibliography: bibliography.bib
---

```{r set-up, include = FALSE, message = FALSE}
library(here)
library(knitr)
library(officedown)
library(officer)
library(tidyverse)
library(flextable)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  dpi = 300,
  fig.cap = FALSE
  )
set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  text.align = "center"
  )
ft <- officer::fp_text(shading.color = "yellow")

word_spec <- function(x, prop = ft) ftext(text = toString(x) ,prop =  ft)
```

Restoration Ecology, TUM School of Life Sciences, Technical University of Munich, Germany

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

\clearpage

\newpage

# Abstract {.unnumbered}

1.  Sowing is a well-established restoration technique to overcome dispersal limitation. Though, adapted seed mixtures to certain site conditions promise a more effective restoration. This is especially important if the restored vegetation has to protect constructions like dikes. An ideal seed mixture-substrate combination would concurrently improve the ecosystem function erosion protection and enhance species-rich grasslands by a fast and successful establishment of target species.
2.  An experiment with 288 plots was conducted on both expositions of a dike and was surveyed 2018--2021. We tested different sand admixtures, substrate depths, seed densities and seed mixture types (hay meadows, dry calcareous grasslands), and measured the persistence of sown species, the Favorable Conservation Status (FCS) and the successional distance to reference sites as calculated from an ordination.
3.  Sand admixture had a significant but slightly positive effect on all indices. Substrate depth, seed density and seed mixture type had no significant effect on persistence or FCS.
4.  On the north exposition, hay meadow-seed mixtures developed closer to its reference than dry grasslands. Overall, the vegetation developed in the desired direction, but a recovery debt remained after four years. Plots still exhibited similarities to negative references from ruderal vegetation.
5.  The south exposition was significantly worse than the north exposition which might be due to the severe drought during establishment and the addition and removal of a geotextile. This initial negative effect remained over the whole observation period.
6.  Synthesis and applications: Adapted seed mixtures can make restorations more efficient like applying hay meadows on the north exposition. However, varying seed mixture-substrate combinations might foster biodiversity like beta diversity more than a desired 'ideal' combination. Furthermore, additional efforts in management might be necessary to reduce the recovery debt and resowing after a severe drought might be sensible due to strong legacy effects.

\clearpage

\newpage

# Keywords {.unnumbered}

artificial soil mixture

community assembly

dry grasslands

Favorable Conservation Status

hay meadow

levee

persistence

river embankment

soil restoration

sowing

species composition

species-rich grassland

\clearpage

\newpage

# Introduction

Grasslands provide several ecosystem services and a high biodiversity [@Dengler.2014; @Bardgett.2021]. However, they are globally endangered [@Bardgett.2021] even by restoration through unsuited afforestation [@dudley2020]. In Europe, calcareous grasslands and hay meadows are on the red list [Category 3, vulnerable, @Janssen.2016]. Restoration is seen as a key factor to sustain ecosystem services and biodiversity [@COP.2014; @UnitedNations.2019b] and sowing is a long-established restoration approach for species-rich grasslands [@Kiehl.2010]. Sowing high-diversity seed mixtures of local provenance produced by commercial but specialized seed companies is a promising way to scale up restoration efforts [@Freitag.2021]. Though, there are still open questions about specifying the seed mixtures to specific site conditions or even more to future climate conditions to improve efficiency and effectiveness of restoration efforts [@Torok.2021].

Restoration ecology can increase predictability of restoration approaches [@Suding.2011; @Mouquet.2015] by using rigorous, repeatable, and transparent experiments linked to a hypothesis derived from theory, which will finally lead to evidence-based restoration [@Cooke.2018; @Wainwright.2018]. Local site conditions and the restoration approach are key predictors of the development of vegetation after sowing [@Brudvig.2017b]. Sowing in general is a way to overcome the dispersal filter [@myers2009; @orrock2023], but we focus on habitat and biotic filtering which are the main assembly processes which can be manipulated by the choice of the right seed mixture-substrate combination [@Torok.2017]. This means a mutually adaption of the substrate to the preferences of the target vegetation, and the seed mixtures to the characteristics of the chosen substrate. The right substrates should reduce habitat filtering and the right seed mixture should minimize competitive exclusion of unsuited species. Modifying seed mixtures for certain site conditions emerged with the rise of functional plant traits [@Funk.2008; @Laughlin.2014; @Balazs.2020] but it is not easy to put it into practice [@merchant2022]. Substrates are artificial soil mixtures and are used in urban areas [@Bauer.2022], quarries [@chenot-lescure2022], or dikes [@Liebrand.1996].

Dikes are promising sites for the restoration of species-rich grasslands because they can increase habitat area and connectivity of semi-natural grasslands and therefore significantly contribute to biodiversity conservation in an agricultural landscape [@Batori.2020]. Dikes have specific characteristics with strongly exposed slopes and river and land orientation [@Batori.2016] and need a vegetation cover to ensure dike stability [@berendse2015]. Dikes can reconcile several requirements [@teixeira2022], which can be fostered by an adapted seed mixture-substrate combination.

The aim of this study is to identify the ideal combination of seed mixtures and substrates for north and south exposed slopes in Central Europe, which develops a vital and species-rich vegetation cover. We set up an experiment and tested agricultural soil with different amounts of sand admixture, substrate depths, seed densities, and seed mixture types. We expected a better development of dry grassland on the south exposition with shallow sandy substrates and hay meadows on the north exposition and less sandy, deep substrates. Furthermore, higher seed densities are recommended for slopes [@KleberLerchbaumer.2017], but the benefit has not been proven, yet.

To measure the success of restoration, we chose four indices. First, we observed the persistence of the sown species, which is a key interest of sowing as a restoration approach [@Wilsey.2021]. For this purpose, the sown species of a specific plot were surveyed, if they are present in that plot in the following years. Second, we calculated the Favorable Conservation Status (FCS) which distinguishes characteristic diversity and derived diversity [@Helm.2015]. Characteristic diversity is created by established species which belong to the habitat-specific species pool. In other words, these species are target species for a habitat type of interest. Third, the successional distance to the reference was determined. This measure quantifies the distance of a plot to the average reference site derived from an ordination [@Rydgren.2019]. To complement these indices, the success of restoration or the difference from desired conditions was evaluated [cf. @Brudvig.2017b]. This was reached by comparing the species composition of the experimental plots with reference sites. We asked following question to evaluate the restoration approaches and the success after four years:

1.  How strong differs the persistence of the sown species among the seed mixture-substrate combinations?

2.  How strong differs the Favorable Conservation Status (FCS) among the seed mixture-substrate combinations?

3.  How close is the established vegetation to the reference state?

\clearpage

\newpage

# Material & Methods

## Experimental design {.unnumbered}

The seed mixture-substrate combinations were tested with an experiment on a dike of the River Danube in SE Germany (Figure \@ref(fig:map); 314 m a.s.l.; WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and an annual precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@hari2020; @cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [Appendix A1, @lfu2021waterlevel; @lfu2021discharge]. For the substrates, sand (0--4 mm) and agricultural soil were used and the soil was from a dike construction site nearby (Steinkirchen). A big roller mixed both components and an excavator put the substrates in the prepared plots.

The target vegetation were lowland hay meadows and semi-dry calcareous grassland [EUNIS code: R22, R1A, @Chytry.2020; Arrhenatherion elatioris, Cirsio-Brachypodion pinnati; EuroVegChecklist: CM01A, DA01B, @mucina2016]. The species pool of hay meadows and dry grasslands consisted of 55 and 58 species, respectively. The seeds were received from a commercial producer of autochthonous seeds. From these species pools, 20 species were randomly selected for each plot (Appendix A2). The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Figure \@ref(fig:traits)). As expected, the hay-meadow mixtures had a higher specific leaf area (SLA) lower seed mass and higher canopy height than the dry-meadow mixtures. The south exposed plots were sown in mid-April 2018 and the north exposition 14 days later. Due to the drought, the south exposition was protected by a geotextile consisting of straw chaff (350 g m^-2^) which was removed after two weeks. The management started with a cut without hay removal and a mowing height of 20 cm in August of the same year (2018), followed by standard deep cuts with hay removal in July 2019 and 2020. In October 2018, *Bromus hordeaceus* was sown as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were halfway up the slopes of the dike, distributed over the north and south exposition, and arranged in six blocks (= replicates). The experiment used a split-plot design combined with a randomized complete block design (Figure \@ref(fig:map)). The split plot was created by the two expositions of the dike, where all 24 treatment combinations were tested: sand admixtures (0%, 25% and 50%), soil depths (15 vs. 30 cm) of practical interest [@KleberLerchbaumer.2017], and the two seed mixture types with the normal seed density and an increased density recommended for slopes [4 vs. 8 g m^-2^ @Kiehl.2010; @KleberLerchbaumer.2017].

Under the substrate, a 5 cm thick drainage layer of gravel (0-16 mm) was installed. Soil samples of the three substrates were tested on both expositions by mixing several sub-samples from different plots. The sand admixture changed the soil texture, increased the C/N ratio, reduced Calciumcarbonat, but did hardly changed the quite high pH values (Table \@ref(tab:substrate)). Phosphate and potassium were rather scarce for agricultural soils and Magnesium rather high [@lfl2022]. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. The establishment rates of species are recorded in Appendix A3. `r word_spec("(Add mean values, range, how many of species pool established, etc.)")` <!--# add more infos -->

## External data

Reference vegetation surveys were extracted from sPlotOpen [@Sabatini.2021] and our own surveys on Danube dikes in the surroundings [@bauer2023]. We selected six calcareous grassland plots [EUNIS code R1A, @Chytry.2020] within SE Germany from sPlotOpen and 98 plots of our own survey which included also lowland hay meadows (R22) and as a negative reference ruderal, dry and anthropogenic vegetation (V38). The functional traits of specific leaf area (SLA), seed mass and canopy height (available traits of all species: 99.2%, 98.8%, 99.6%) were retrieved from the databases LEDA [@Kleyer.2008] and TRY [@Kattge.2020].

## Statistical analyses {.unnumbered}

To measure the effects of the treatments on our three response variables, we calculated Bayesian linear mixed-effects models (BLMM) with the random effect plot nested in block with the Cauchy prior [see @lemoine2019]. Furthermore, we included as a fixed effect the botanists of each plot. For the simple effects of the treatments, we chose plausible weakly informative priors. To evaluate the influence of the priors, prior predictive checks and models with non-informative priors were calculated. For the computation, we used 4 chains, a thinning rate of 2, 5000 iterations for warm-up and 10,000 in total. For evaluating the computation, MCMC chain convergence was checked with trace plots and by the R-hat values, and MCMC chain resolution by the effective sampling size (ESS). Posterior predictive checks were done with Kernel density estimates histograms of statistics skew and leave-one-out (LOO) cross-validation [see @gabry2019]. Finally, models were compared with the Bayes factor (BF) and Bayesian *R*² values [@gelman2019]. A non-metric multidimensional scaling ordination (NMDS) with Jaccard dissimilarity was used to visualize variation in species composition in space and time (2D-stress 0.21). Seven species were excluded because they had an accumulated cover over all plots of less than 0.5%.

Data, code and the whole model specification and evaluation is stored on Zenodo and in an easily accessible scroll-down document on GitHub `r word_spec("(include Zenodo reference)")`.<!--# Cite Zenodo--> There, the sections are referenced to the Bayesian analysis reporting guidelines [BARG, @kruschke2021]. All analyses were performed in R [Version 4.2.2, @base] with the functions 'brm' from the package 'brms' [@brms] for model calculation, several functions from 'brms' and 'bayesplot' [@bayesplot] for model evaluation, and 'metaMDS' from 'vegan' for the ordination [@vegan].

\clearpage

\newpage

# Results

## Weak effects of restoration treatments

We could identify a statistically clear positive effect of sand admixture on the persistence of sown species and on the recovery rate (Figure \@ref(fig:effects)). The posterior distributions are also shown in the interaction plots which separate exposition and survey year indicate. For all three response variables, the vegetation developed strongly positive after one year, but the recovery rate slowed down in the following years (Figures \@ref(fig:persistence), \@ref(fig:fcs), \@ref(fig:recovery)). Both expositions have similar trends but for all responses the values are clearly lower on the south exposition, e.g., persistence values are on average more than 46% higher on the north exposition (Figure \@ref(fig:persistence)). In the interaction plots, the effects of restoration treatments are neither clear nor strong (Figures \@ref(fig:persistence), \@ref(fig:fcs), \@ref(fig:recovery)). Persistence of both seed mixture types seems to be slightly positive affected by sand admixture on the north exposition (+ 6--7 ± 4%, Figure \@ref(fig:persistence)).

## Hay meadows on north exposition closer to reference

The ordination visualizes the species composition of seed mixtures and the development of the plots over time (Figure \@ref(fig:nmds)). The NMDS confirms that the seed mixtures are variable, but distinctive for hay meadows and dry grasslands and ascertain the right direction of the vegetation development. One exception is that the development of intended hay meadows on the south exposition does not develop towards its seed-mixture compositions.

The ordination presents the experimental plots also in context with reference sites. The reference sites have a bigger variation than the seed mixtures and are close to the seed mixtures but hardly overlap. The positions of reference sites are shifted to the left in comparison to the seed mixtures which means in the direction of early successional states. Though, they are still different from the negative references of ruderal, dry, anthropogenic vegetation. Negative references were only available on the south exposition and they are located in the NMDS between the positive reference sites and the state of restored plots in 2021. Hay meadow-seed mixtures led to a closer development to hay meadow references than dry grasslands to their references (Figure \@ref(fig:effects), \@ref(fig:recovery)) this is especially the case on the north exposition (Figure \@ref(fig:nmds)).

\clearpage

\newpage

# Discussion

The vegetation of the dike experiment developed in the right direction, but with reduced velocity and significantly and lasting worse development of the south exposition. Sand admixture of 25% was sufficient to reach a positive effect. All in all, it was hardly possible to identify an ideal seed mixture-substrate combination. Though, hay meadows developed closer to its reference on the north exposition than dry grasslands.

## General effects of treatments and exposition

As expected, sand admixture led to higher persistence of sown species but only on the north exposition and the effect size of about 6% in the fourth year after restoration is rather small. In addition, 50% admixture did not further improve the result. Restoration on agricultural soils can have limited success due to high nutrient loads [@walker2004] but can lead to positive results, if mixed with mineral material as sand and gravel [@chenot-lescure2022]. Substrate depth did not significantly affect persistence or Favorable Conservation Status (FCS) similarly to earlier studies [@baer2004]. Larger differences of soil depth might be necessary to create negative effects by greater soil depths as was shown for prairies [@dornbush2010]. Neither, seed density had an effect on persistence and FCS which fits to result of @Kaulfuss.2022, who found out that a certain amount of seeds are necessary but higher densities do not further improve establishment of target species but have rather a slight negative effect.

The vegetation on the south exposition performed consistently and significantly worse than the north exposition and had a different species composition which was also shown by @Batori.2016 in Hungary. Here, the differences might have some unwanted experimental reasons: the geotextile, which had been implemented on the south exposition, was removed after two weeks because of the death of several protected lizards. This led to the drying of many seedlings why the distinction of the seed mixture types might be leveled on the south exposition. Moreover, the intense drought of 2018 and 2019 [@hari2020] might have decreased germination and establishment of sown species because a moist spring with a mesic temperature improves establishment [@Larson.2021; @orrock2023]. The interesting point here is the lasting negative effect which suggests a legacy effect by unfavorable weather conditions after sowing as it was observed by other studies [@Stuble.2017; @Groves.2020]. The problems during the establishment phase might have led to an alternative trajectory and could probably only be met by resowing.

## No interaction effect of seed mixture-substrate combinations

Our aim was to identify a ideal seed mixture-substrate combination in respect of restoration effectiveness and biodiversity. For evaluating effectiveness, we measured the establishment and persistence of the sown species and for fostering plant biodiversity in general, we measured the FCS. Though, we could not identify an interaction effect for neither of the indices. We would have expected a better performance of hay-meadow seed mixtures with lower sand admixture and for dry grasslands with higher sand admixture. Our results suggest that, at least after four years, the substrate conditions are within the range of both seed mixture types (hay meadows, dry grasslands). Although, both types are clearly plant-sociologically and functionally distinct, they are still relative closely related why they contain shared species and could develop on similar site conditions [Figure @ref(fig:traits); @Oberdorfer.1993; @Oberdorfer.1993b]. Other grassland studies could identify more or less clear interactions of opposing habitat preferences or functional traits along the gradients of productivity, moisture, and nutrients [@Zirbel.2020; @Freitag.2021; @Kaulfuss.2022]. However, these two studies did not work with an experimental set up of different seed mixture-substrate combinations, but analyzed the result of habitat and biotic filtering after five and about 15 years, respectively. Furthermore, the non-existence of an ideal combination could be explained by priority effects which means that the species of the sown species could establish earlier than others and preempted the available niches [@Fukami.2015].

## The success of the restoration approaches

The seed mixtures and positive references were similar but hardly overlapped (Figure \@ref(fig:nmds)). The position on the NMDS suggests that the seed mixture represent a latter successional stage than the positive references. The NMDS shows a slightly better adaption of hay meadows to the north exposition than of dry grasslands (Figure @ref(fig:recovery)). This can be expected from the requirements of hay meadows for mesic conditions which can be provided on north exposed dike slopes [@Batori.2020; @Oberdorfer.1993b]. On the south exposition, the plots of hay meadows developed rather to dry grassland references which indicates an inefficient restoration by using an non-adapted seed mixture.

The vegetation developed in general in the right direction but was still distinct from positive references and seed mixtures after four years. It was rather similar to the negative reference of perennial ruderal vegetation. The gap between goal and restoration outcome was also shown for other sowing experiments or restorations [@mitchley2012; @Engst.2016; @Kaulfuss.2022] or for dike vegetation compared with semi-natural references [@Batori.2016]. This result is not surprising since the 'recovery debt' is a general phenomenon of grassland restoration [@MorenoMateos.2017; @jones2018].

\clearpage

\newpage

# Conclusions

Our results suggest that adapted seed mixtures can increase restoration efficiency by sowing hay meadows on the north exposition and not on the south exposition of dikes. Furthermore, to reduce nutrient load by sand admixture has a clear positive effect, although the effect is small. The question remains if money is best invested in sand admixture or in other restoration measures. It seems not to be necessary to increase seed density on dike slopes, which would reduce restoration costs. As well as the use of more deeper soil depth which would otherwise create costs for removal of topsoil. There was no perfect seed mixture-substrate combination, why we conclude that a variation of seed mixtures types and different soils along restoration sections would promote biodiversity better than a single solution [@bauer2023; @Holl.2022]. The strong and lasting negative effect of drought on the south exposition might require a resowing. To close the recovery debt, the adaption of the management might be promising since it is a relevant factor aside of restoration approach and site characteristics [@Grman.2013; @tölgyesi2021]. For example, starting sheep grazing on the restored sites will modify the disturbance regime and improve dispersal. All in all, our results support the finding that dike grasslands can promote biodiversity in agricultural landscapes [@Batori.2020]. Though, the recovery debt highlights that restored grasslands cannot substitute old-growth grasslands [@nerlekar2020; @Buisson.2022].

\clearpage

\newpage

# Acknowledgements {.unnumbered}

We would like to thank our project partners Dr. Markus Fischer, Frank Schuster, and Christoph Schwahn (WIGES GmbH) and Klaus Rachl and Stefan Radlmair (Regierung von Niederbayern) for numerous discussions on restoration and management of dike grasslands. Field work was supported by Clemens Berger and Uwe Kleber-Lerchbaumer (Wasserwirtschaftsamt Deggendorf). We thank Holger Paetsch, Simon Reith, Anna Ritter, Jakob Strak, Leonardo H. Teixeira, and Linda Weggler for assisting with the field surveys or soil analyses in 2018--2020. The German Federal Environmental Foundation (DBU) supported MB with a doctoral scholarship. The study has been supported by the TRY initiative on plant traits ([http://www.try-db.org).](http://www.try-db.org).) The TRY initiative and database is hosted, developed and maintained by J. Kattge and G. Boenisch (Max Planck Institute for Biogeochemistry, Jena, Germany). TRY is currently supported by Future Earth/bioDISCOVERY and the German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig.

# Author contribution {.unnumbered}

JH and JK designed the experiment. JH did the surveys in the years 2018--2020 and MB in 2019 and 2021. MB did the analyses and wrote the manuscript. JK and JH critically revised the manuscript.

# Open research {#sec-open-research .unnumbered}

Data and code are stored on Zenodo (Bauer et al. 2023) <!--# Update citation after upload -->

Model evaluation is stored on GitHub: [github.com/markus1bauer/2023_danube_dike_experiment](https://github.com/markus1bauer/2023_danube_dike_experiment/tree/main/markdown)

# Funding {.unnumbered}

MB is funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020 (No. 80 002 312).

\clearpage

\newpage

# References {.unnumbered}

::: {#refs}
:::

\clearpage

\newpage

<!---BLOCK_LANDSCAPE_START--->

# Tables {.unnumbered}

## Table 1 {.unnumbered}

```{r tab.id = "seedmix", ft.align = "left", tab.cap = "Each plot received an individual set of 20 species with some restrictions to the number of species per functional group. The total species pool for hay meadows was 55 and for dry grassland 58. In total, plots received 16 or 32 g of seeds. All individual seed mixtures are stored in Appendix A3."}
read_csv(here::here("data", "raw", "data_raw_table_1_seedmix.csv"),
         show_col_types = FALSE) %>%
  slice(-1) %>%
  flextable() %>%
  merge_at(part = "header", j = 2:3) %>%
  align(align = "center", part = "all", j = -1) %>%
  add_header_row(top = FALSE, values = c("", "Hay meadow", "Dry grassland", "", "", "")) %>%
  add_header_row(top = FALSE, values = c("", "#", "#", "#", "wt%", "wt%")) %>%
  bold(part = "header", i = 1) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  autofit()
```

\clearpage

\newpage

## Table 2<!--# Check quotation --> {.unnumbered}

```{r tab.id = "substrate", ft.align = "left", tab.cap = "Soil of the three substrates was analyzed for the fraction < 2 mm. Soil texture was classified according to the 'Bodenkundliche Kartieranleitung' [@BodenkundlKartieranl.2005]. The pH was measured in CaCl2 solution. Plant available phosphorus and potassium were measured in a calcium acetate-lactate extract and magnesium in CaCl2 extract. For calculating CaCO3, a subsample was annealed at 550 °C and the measured C amount calculated with 8.33. To calculate total N and the C/N ratio, a subsample was incinerated at 1000 °C. Lt3 = medium clayey loam; Ls4 = strong sandy loam; Sl3 = medium loamy sand; Sl4 = strong loamy sand."}
read_csv(here::here("data", "raw", "data_raw_table_3_substrate.csv"),
         show_col_types = FALSE) %>%
  select(-starts_with("lfl")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(
      here::here("data", "raw", "data_raw_table_3_substrate.csv"),
                 show_col_types = FALSE) %>%
      select(-starts_with("lfl")) %>%
      slice(1)
    ) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  align(align = "center", part = "all", j = -1) %>%
  compose(i = 1, j = 10, part = "header", value = as_paragraph("P", as_sub("2"), "O", as_sub("5"))) %>%
  compose(i = 1, j = 11, part = "header", value = as_paragraph("K", as_sub("2"), "O")) %>%
  compose(i = 1, j = 12, part = "header", value = as_paragraph("Mg", as_sup("2+"))) %>%
  compose(i = 1, j = 14, part = "header", value = as_paragraph("CaCO", as_sub("3"))) %>%
  compose(i = 2, j = 10, part = "header", value = as_paragraph("mg 100 g", as_sup("−1"))) %>%
  compose(i = 2, j = 11, part = "header", value = as_paragraph("mg 100 g", as_sup("−1"))) %>%
  compose(i = 2, j = 12, part = "header", value = as_paragraph("mg 100 g", as_sup("−1"))) %>%
  align(align = "left", part = "footer") %>%
  set_table_properties(layout = "autofit", width = 1)
```

\newpage

<!---BLOCK_LANDSCAPE_STOP--->

# Figures {.unnumbered}

## Figure 1 {.unnumbered}

```{r fig.id = "map", fig.asp = 1.2854, fig.width = 5, fig.cap = "The experiment was located on a dike at river Danube in SE Germany.  The 288 plots were allocated in six blocks (red squares on upper photo) and on the north and south exposition [central photo; both aerial photos: @bayerischevermessungsverwaltung2023]. Four treatments were conducted (Sand admixture, substrate depth, seed density, seed mixture types H and D). The western half of a block was with a shallow substrate depth for the blocks and within this half the substrates. The photo on the bottom shows one block from the side after four years (Photo: Markus Bauer)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_1_map.png")
  )
```

\clearpage

\newpage

## Figure 2 {.unnumbered}

```{r fig.id = "traits", fig.asp = 0.3030, fig.width = 6.49, fig.cap = "The community-weighted means (CWM) of the seed mixtures used for the field experiment of the functional traits specific leaf area (SLA), seed mass and canopy height. The CWMs were weighted by seed weight in the seed mixture and  indicate differences between the target types hay meadows and dry grasslands . "}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm_traits.tiff")
  )
```

\clearpage

\newpage

## Figure 3

```{r fig.id = "effects", fig.asp = 0.3030, fig.width = 6.49, fig.cap = "The posterior density distributions (grey) of the treatments calculated over all four surveyed years. Furthermore, the median and the 66% and 95% credible intervals are shown. We evaluated the persistence of sown species (A), the Favorable Conservation Status (FCS; B) and the successional distance to reference (recovery; C). The FCS is the ratio of target species and non-target species. Note that the zero lines indicate that both levels have equal values. This means that hay meadows are closer to their reference than dry grasslands  (C)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff")
  )
```

\clearpage

\newpage

## Figure 4 {.unnumbered}

```{r fig.id = "persistence", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "Persistence of sown species on a dike experiment over the years. The substrates had different sand admixtures and different seed mixture types were sown. Per plot were 20 species sown. Shown are the posterior distributions, their medians and 95 % credible intervals derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x8cm.tiff")
  )
```

\clearpage

\newpage

## Figure 5 {.unnumbered}

```{r fig.id = "fcs", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The Favorable Conservation Status (FCS) of a dike experiment. The substrates had different sand admixtures and different seed mixture types were sown. The zero line indicate that target and non-target species are balanced. Positive values indicate that there are more target species.  Shown are the posterior distributions, their medians and 95 % credible intervals derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x8cm.tiff")
  )
```

\clearpage

\newpage

## Figure 6 {.unnumbered}

```{r fig.id = "recovery", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The successional distance (*d~jt,0~*) of plots of an experiment on a dike to its respective reference sites (hay meadows or dry grasslands). The substrates had different sand admixtures and different seed mixture types were sown. The zero lines indicate the mean position of the reference sites on the NMDS axis 1 (Figure 7). The grey area marks the standard deviation of the position of the reference sites. Shown are the posterior distributions, their medians and 95 % credible intervals derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x8cm.tiff")
  )
```

\clearpage

\newpage

## Figure 7 {.unnumbered}

```{r fig.id = "nmds", fig.asp = 0.9696, fig.width = 6.49, fig.cap = "NMDS ordination based on Sørensen dissimilarity of the species composition of 288 plots on dike grasslands over four years after sowing the seed mixtures (black squares).  In addition, 98 positive and negative reference plots were included (filled symbols) from older dike grasslands in the surroundings [@bauer2023] and six plots from sPlotOpen [@Sabatini.2021].  The circles show the standard error of the groups. 2D stress: 0.21."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff")
  )
```

\clearpage

\newpage

## Session Info {.unnumbered}

```{r sessioninfo}
sessionInfo()
```
