---
title: "Dike experiment"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  bookdown::word_document2:
    reference_docx: StylesTemplate.docx
    global_numbering: true
csl: template-journal-of-vegetation-science.csl
bibliography: bibliography.bib
---

# Document head {.unnumbered}

Restoration Ecology, TUM School of Life Sciences, Technical University of Munich, Germany

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

Open Research: Data and code are permanently available on Zenodo under: <!--# https://doi.org/10.XXXX -->.

**Max 8000 words**

\newpage

```{r set-up, include = FALSE}
library(knitr)
library(bookdown)
library(officedown)
library(officer)
library(tidyverse)
library(flextable)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  dpi = 50
  )
set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 10
  )


# read analysis
#rmarkdown::render(here::here("results", "scripts", "Models.Rmd"), quiet = TRUE) # quiet to not get text from rendering
```

# Abstract (max 300 words) {.unnumbered}

\clearpage

\newpage

# Keywords {.unnumbered}

Bayesian models

dike grasslands

embankment

endangered species

environmental filter

species composition

restoration

sand

soil properties

surveys

vegetation

\clearpage

\newpage

# Introduction

"Ecological theory can point restoration toward important processes that need manipulation [...]. However, for this information to be relevant, restoration ecology needs to employ evidence-based assessments [...]" (Suding 2011) --\> No evidence for theory with our project

-   Endangered grasslands

-   Predictability of restorations --\> experiments, evidence-based restoration (cooke 2018)

-   Brudvig 2017 among restoration outcomes and in comparison to reference

-   Substrate modification to influence habitat filtering

-   Sowing a typical approach + priority effects + species pool

-   Introduce FCS, Recovery time, Persistence

-   Dike grasslands as case study

-   The aim of the study is

We asked following question to evaluate the restoration success after four years:

1.  How strong differs the persistence of the sown species among the seed mixture-substrate combinations?

2.  How strong differs the Favourable Conservation Status (FCS) among the seed mixture-substrate combinations?

3.  How close is the vegetation to the reference state?

\clearpage

\newpage

# Material & Methods

## Experimental design {.unnumbered}

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:figure-map)), 314 m above sea level (asl); WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [@lfu2021waterlevel,@lfu2021discharge] (Table \@ref(fig:table-flood-weather)). <!--# table ref is not working --> Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from the dike <!--# Is that right?  --> and a with an excavator the substrates were mixed <!--# Is that right?  --> and the plots prepared. The target vegetation were lowland hay meadows (Arrhenatherion elatioris, CM01A) and calcareous grassland (Cirsio-Brachypodion pinnati, DA01B code of the EuroVegChecklist, @mucina2016 ). The species pool of hay meadows and calcareous grasslands consisted of 55 and 58 species <!--# check numbers  -->, respectively. The seeds were received from a commercial producer of autochtonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(fig:table-seedmix))<!--# check  -->. The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(fig:table-3)) <!--# More details and check  -->. The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, Bromus hordeaceus was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-map)). The split plot was caused by testing all 24 treatments on both sides of the dike. We tested three substrates with 0%, 25% and 50% sand admixture and two soil depths (15 vs. 30 cm). Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table \@ref(fig:table-substrate)). The two seed mixture types were tested with two seeding densities (4 vs. 8 g m\^-2 ) <!--# include superscript -->. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. Establishment rates of species <!--# Refer to Table Sx -->.

<!--# We measured FCS, recovery, persistence -->

<!--# Mention traits -->

## Statistical analyses {.unnumbered}

A non-metric multidimensional scaling ordination (NMDS) with Jaccard dissimilarity was used to visualize variation in species composition in space and time. Rare species with an accumulated cover over all plots of less than 0.5% were excluded (2D-stress: 0.21).

To measure the effects of the treatments on our three response variables, we calculated Bayesian linear mixed effects models (BLMM) with the random effect plot nested in block with the Cauchy prior [see @lemoine2019]. Furthermore, we included as a fixed effect the botanists of each plot. For the simple effects of the treatments we chose plausible weakly informative priors. To evaluate the influence of the priors, prior predictive checks and models with non-informative priors were calculated. For the computation, we used 4 chains, a thinning rate of 2, 5,000 iterations for warm-up and 10,000 in total. For evaluating the computation, MCMC chain convergence was checked with trace plots and by the R-hat values and MCMC chain resolution by the effective sampling size (ESS). Posterior predictive check was done with Kernel density estimates histograms of statistics skew and leave-one-out (LOO) cross-validation [see @gabry2019]. Finally, models were compared with the Bayes factor (BF) and Bayesian *R*² values [@gelman2019].

The software and the whole model specification and evaluation is stored on Zenodo <!--# insert citation --> and in an easily accessible a scroll-down document on GitHub (see \@ref(open-research)). There, the sections are referenced to the Bayesian analysis reporting guidelines [BARG, @kruschke2021].

All analyses were performed in R [Version 4.2.2; @base] with the functions 'brm' from the package 'brms' [@brms] for model calculation, several functions from 'brms' and 'bayesplot' [@bayesplot] for model evaluation and 'metaMDS' from 'vegan' for the ordination [@vegan].

<mark>To identify the final model, we first reviewed the residual diagnostics of the candidate models and subsequently compared the remaining models using the Akaike information criterion adjusted for a small sample size (AICc) and chose the most parsimonious model (Appendix A4). Finally, we calculated the marginal and conditional coefficients of determination (R²m, R²c) and the 95% confidence intervals of the response variables.</mark>

\clearpage

\newpage

# Results

We could identify a statistically clear positive effect of sand addition on the persistence of sown species and on the recovery rate (Figure \@ref(fig:figure-effects)). Furthermore, sown hay meadows were clearly closer to their reference plots than sown dry grasslands were to their references (Figure \@ref(fig:figure-effects)).

The posterior distributions shown in the interactions plots which separates exposition and survey year is less clear. <!--# Go on here -->

\clearpage

\newpage

# Discussion

Mention drought in 2018 and 2019

\clearpage

\newpage

# Conclusions

Text

\clearpage

\newpage

# Acknowledgements {.unnumbered}

We would like to thank our project partners Dr. Markus Fischer, Frank Schuster, and Christoph Schwahn (WIGES GmbH) and Klaus Rachl and Stefan Radlmair (Regierung von Niederbayern) for numerous discussions on restoration and management of dike grasslands. Field work was supported by Clemens Berger and Uwe Kleber-Lerchbaumer (Wasserwirtschaftsamt Deggendorf). We thank Holger Paetsch, Simon Reith, Anna Ritter, Jakob Strak, Leonardo H. Teixeira, and Linda Weggler for assisting with the field surveys or soil analyses in 2018--2020. The German Federal Environmental Foundation (DBU) supported MB with a doctoral scholarship.

# Author contribution {.unnumbered}

JH and JK designed the experiment. JH did the surveys in the years 2018--2020 and MB in 2019 and 2021. MB did the analyses and wrote the manuscript. JK and JH critically revised the manuscript.

# Open research {.unnumbered}

Data and code is available on Zenodo: <!--# https://doi.org/10.XXXX -->

Model evaluation is stored on GitHub: [github.com/markus1bauer/2023_danube_dike_experiment](https://github.com/markus1bauer/2023_danube_dike_experiment/tree/main/markdown)

# Funding {.unnumbered}

MB is funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020. <!--# funding number of WIGES -->

# References {.unnumbered}

::: {#refs}
:::

\clearpage

\newpage

<!---BLOCK_LANDSCAPE_START--->

# Tables {.unnumbered}

## Table 1

Mischprobe aus mehreren Plots. Measured in \< 2mm fraction <!--# Write full description -->

```{r table-substrate, fig.cap = "table_substrate"}
read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>% slice(1)
    ) %>%
  autofit()
```

\clearpage

\newpage

## Table 2 {.unnumbered}

Each plot received a individual set of 20 species with some restrictions to the number of species per functional group. <!--# Full description -->

```{r table-seedmix, fig.cap = "table_seedmix"}
read_csv(here::here("data", "raw", "data_raw_table_2.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(top = FALSE, values = c("", "", "#", "#", "%", "%")) %>%
  autofit()
```

\clearpage

\newpage

## Table 3

functional traits <!--# full description + table -->

```{r table-traits, fig.cap = "table_traits"}
read_csv(here::here("data", "raw", "data_raw_table_2.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  autofit()
```

<!---BLOCK_LANDSCAPE_STOP--->

\newpage

# Figures {.unnumbered}

## Figure 1 {.unnumbered}

```{r figure-map, fig.cap = "figure-map"}
knitr::include_graphics(here::here("outputs", "figures", "figure_1_map_tmap_300dpi_8x11cm.tiff"))
```

\clearpage

\newpage

## Figure 2 {.unnumbered}

```{r figure-effects, fig.cap = "figure-effects"}
knitr::include_graphics(here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff"))
```

\clearpage

\newpage

## Figure 3 {.unnumbered}

```{r figure-persistence, fig.cap = "figure-3"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x14cm.tiff"))
```

\clearpage

\newpage

## Figure 4 {.unnumbered}

```{r figure-fcs, fig.cap = "figure-4"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x14cm.tiff"))
```

\clearpage

\newpage

## Figure 5 {.unnumbered}

```{r figure-recovery, fig.cap = "figure-5"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x14cm.tiff"))
```

Fig. 5:

\clearpage

\newpage

## Figure 6 {.unnumbered}

```{r figure-nmds, fig.cap = "figure-6"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff"))
```

Fig. 6:

\clearpage

\newpage

# Supplementary Material {.unnumbered}

## Table S1 {.unnumbered}

Three dry years (2018--2020) and three minor floods (2018, 2019, 2021) occurred during the study period. Annual temperature, precipitation and discharge of River Danube near the study sites (2017--2021), based on weather station Metten [mean, 1981--2010; ID, 3271; WGS84, lat/lon, 48.85476, 12.918911; @cdc2022prec, @cdc2022temp] and stream gauge Pfelling [ID, 10078000; WGS84, lat/lon, 48.87975, 12.74716; @lfu2021discharge]. HQ2 = Highest discharge with a probability of occurrence every second year. HSW = Highest water level for shipping.

```{r table-flood-weather, fig.cap = "table-s1"}
read_csv(here::here("data", "raw", "data_raw_table_s1.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  compose(part = "body", i = ~ Unit == "m3 s−1", j = "Unit", value = as_paragraph("m", as_sup("3"), " s", as_sup("−1"))) %>%
  autofit()
```

\clearpage

\newpage

## Table S2 {.unnumbered}

Establishment rate <!--# Full description an include it here -->

```{r table-establishment, fig.cap = "table-s2"}
knitr::include_graphics(here::here("outputs", "tables", "table_a2_establishment.html"))
```

\clearpage

\newpage

## Table S3 {.unnumbered}

See repository outputs/tables/table_a3_seedmixes.csv <!--# Write full link and full description -->

\clearpage

\newpage

## Session Info {.unnumbered}

```{r sessioninfo}
sessionInfo()
```
